---
title: "Hungarian Chickenpox Analysis"
output: html_notebook
---

In this notebook there will be an analysis of Hungarian weekly chickenpox cases. First, let's have a look at the data.

```{r}
# Load data
data = read.csv('hungary_chickenpox.csv')
print(class(data))
```
```{r}
print(data)
```
Is possible to see how the dataset contains 522 rows (weeks) and 21 columns (1 for the date and 20 cities).
Now we must convert the first column in the proper format and transform the dataset in a timeseries.
```{r}
data$Date <- as.Date(data$Date, format='%d/%m/%Y')
print(str(data))
```
```{r}
library(zoo)
zoo_data <- zoo(data[2:ncol(data)], order.by=data$Date)
print(str(zoo_data))
```
```{r}
plot.zoo(zoo_data, plot.type='multiple', nc=2,
         main='Hungarian Chickenpox Weekly Cases', yax.flip = TRUE)
```
A seasonal behavior is clear in every city. It seems also that the peaks coincide.Let's check.

```{r}
library(RColorBrewer)
colors <- c(brewer.pal(n=12, name='Set3'), brewer.pal(n=8, name='Dark2'))
plot.zoo(zoo_data, plot.type = 'single',
         main='Hungarian Chickenpox Weekly Cases', col = colors,
         ylab = 'Weekly cases')
```
For a better comparison, let's scale the data in order to eliminate differences in population.
```{r}
min_max_scaler <- function(x){
  (x - min(x)) / (max(x) - min(x))
}
data_scaled <- as.data.frame(lapply(data[, -1], min_max_scaler))
zoo_scaled <- zoo(x = data_scaled, order.by=data$Date)
plot.zoo(zoo_scaled, plot.type='single',
         main='Hungarian Chickenpox Weekly Cases SCALED', col=colors)
```
Let's now compute the correlation matrix.
```{r}
corr_matrix <- cor(zoo_data, method = 'pearson')
library(corrplot)
corrplot(corr_matrix, method = 'square', type='full', addCoef.col = 'black',
         number.cex = 0.5, diag = FALSE, number.digits = 1, tl.cex=0.5, tl.col = 'black')
print(min(corr_matrix))
```
As expected, the correlation is very high between all the cities (always higher than 0.25).































