---
title: "Hungarian Chickenpox Analysis"
output: html_notebook
---

In this notebook there will be an analysis of Hungarian weekly chickenpox cases. First, let's have a look at the data.

# EDA

```{r}
# Load data
data = read.csv('hungary_chickenpox.csv')
print(class(data))
```
```{r}
print(data)
```
Is possible to see how the dataset contains 522 rows (weeks) and 21 columns (1 for the date and 20 cities).
Now we must convert the first column in the proper format and transform the dataset in a timeseries.
```{r}
data$Date <- as.Date(data$Date, format='%d/%m/%Y')
print(str(data))
```
```{r}
library(zoo)
zoo_data <- zoo(data[2:ncol(data)], order.by=data$Date)
#print(str(zoo_data))
plot.zoo(zoo_data, plot.type='multiple', nc=2, xlab = 'Date',
         main='Hungarian Chickenpox Weekly Cases', yax.flip = TRUE)
```
A seasonal behavior is clear in every city. It seems also that the peaks coincide.Let's check. For a better comparison, let's scale the data in order to eliminate differences in population between cities.

```{r}
library(RColorBrewer)
colors <- c(brewer.pal(n=12, name='Set3'), brewer.pal(n=8, name='Dark2'))
min_max_scaler <- function(x){
  (x - min(x)) / (max(x) - min(x))
}
data_scaled <- as.data.frame(lapply(data[, -1], min_max_scaler))
zoo_scaled <- zoo(x = data_scaled, order.by=data$Date)
plot.zoo(zoo_scaled, plot.type='single', xlab = 'Date',
         main='Hungarian Chickenpox Weekly Cases SCALED', col=colors)
```
Let's now compute the correlation matrix.
```{r}
corr_matrix <- cor(zoo_data, method = 'pearson')
library(corrplot)
corrplot(corr_matrix, method = 'square', type='full', addCoef.col = 'black',
         number.cex = 0.5, diag = FALSE, number.digits = 1, tl.cex=0.5, tl.col = 'black')
print(paste('Minimum correlation value is:', min(corr_matrix)))
```
As expected, the correlation is very high between all the cities (always higher than 0.25).

Now, Let's focus on the country aggregate.
```{r}
total_cases <- rowSums(data[, -1])
total_cases <- zoo(total_cases, order.by = data$Date)
plot.zoo(total_cases, main = 'Hungarian Chickenpox Cases (Reported)',
         ylab = 'Count', xlab = 'Date')
```
It is clear a seasonal pattern (more cases in cold months?) and is also possible to notice that peaks' height is decreasing. To reduce oscillations, a monthly aggregation is now performed. There will be used a "sum" aggregation function as datasets refers to reported cases and not active cases.
```{r}
monthly_data <- aggregate(zoo_data, as.yearmon, sum)
monthly_total <- aggregate(total_cases, as.yearmon, sum)
total_ts <- ts(monthly_total, start = c(2005, 1), freq=12)
print(total_ts)
plot(monthly_data, yax.flip = TRUE, main = 'Monthly Reported Cases', xlab = 'Date')
plot(monthly_total, main = 'Monthly Reported Cases', ylab = '', xlab = 'Date')
boxplot(total_ts ~ cycle(total_ts), xlab = 'Months', ylab = 'Cases',
        main='Monthly Cases Distribution')
```
As expected, the number of reported cases is higher in cold month that are also months in which schools are open and virus can spread easily.

# Decomposition

First, we will try with the additive method.
```{r}
cases_decompose <- decompose(total_ts)
plot(cases_decompose)
par(mar = c(5, 4, 6, 2))
plot(acf(c(cases_decompose$random), na.action = na.pass, plot = FALSE),
     main = 'Residuals ACF (Additive)')
```
The downward trend is clear. Is remarkable that the random component's volatility is lower when cases are low; this suggest to use the multiplicative decomposition. We see also a significant autocorrelation when the lag is 4 and 5.
```{r}
cases_decompose <- decompose(total_ts, type = 'multiplicative')
plot(cases_decompose)
par(mar = c(5, 4, 6, 2))
plot(acf(c(cases_decompose$random), na.action = na.pass, plot = FALSE),
     main = 'Residuals ACF (Multiplicative)')
```
With the multiplicative approach, the volatility of the random component is more constant and the autocorrelation in the random component is less significant (but we still have a significant autocorrelation with lag = 4).

# Forecasting

# Model Evaluation

# Time Frequency Analysis

# ARCH / GARCH

# VAR




























